---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.path = "man/figures/",
  dpi = 300
)
```

# odes


This package implements an approach to forecasting count time series with a form of exponential smoothing built from observation-driven models of Davis, Dunsmuir and Streett (2003). The methodology is discussed in Karlis, Pedeli and Varin (2023).


<!-- badges: start -->
<!-- badges: end -->

 

## Installation

You can install the development version of odes from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("crisvarin/odes", build_vignettes = TRUE)
```

## Example

This is a basic example showing how to use odes for disease surveillance:
```{r example}
library(odes)
example(surveillance.odes)
```

The package vignette expands on the previous example by allowing the results discussed in Section 4 of Karlis, Pedeli, and Varin (2023) to be fully 
replicated:
```{r}
# To view the vignette type
# vignette("odes")
```



## References

Davis, R.A., Dunsmuir, W.T. & Streett, S.B. (2003). Observation-driven models for Poisson counts. *Biometrika* **90**(4), 777–790.

Karlis, D., Pedeli, X. & Varin, C. (2023). Observation-driven exponential smoothing. *Stat* **12**(1), e642. 
