---
title: "Observation-driven Exponential Smoothing for Poisson Counts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Observation-driven Exponential Smoothing for Poisson Counts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 4, fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to replicate step-by-step the analysis of Dengue fever 
in Italy discussed in Section 4 of Karlis, Pedeli and Varin (2023) using the 
`odes` package. 

Let us load `odes`:
```{r setup}
library("odes")
```

The next step is to create a data frame with dengue fever data. The data frame 
columns are `counts` with Dengue cases and the `time` index:
```{r}
dengue_data <- data.frame(counts = dengue, time = seq_len(length(dengue)))
```

It is also convenient to name the rows of the data frame with the observation 
dates:
```{r}
dates <- seq(as.Date("2008/1/1"), by = "month", length.out = nrow(dengue_data))
rownames(dengue_data) <- dates
```

We now split the data into the phase I subset that will be used for model fitting 
and the phase II subset that will be used for surveillance:
```{r}
dengue_phaseI <- dengue_data[1:108, ]
dengue_phaseII <- dengue_data[-(1:108), ]
```


We can now fit ODES to the phase I data:
```{r}
fit <- odes(counts ~ sin(2 * pi * time / 12) + cos(2 * pi * time / 12),
data = dengue_phaseI)
summary(fit)
```

Finally, surveillance of phase II data:
```{r}
surv <- surveillance.odes(fit, new_data = dengue_phaseII, names_as_dates = TRUE, 
                          date_labels = "%Y", breaks = "2 year")
```

The allarms identified  by `odes` are: 
```{r}
surv$alarms
```

For comparison, the following lines repeat the previous analysis but without resorting to robustness adjustment (`Huber = FALSE`):
```{r}
fit2 <- update(fit, Huber = FALSE)
summary(fit2)
surv2 <- surveillance.odes(fit2, new_data = dengue_phaseII, 
                           names_as_dates = TRUE, date_labels = "%Y", 
                           breaks = "2 year", x_label = NULL)
surv2$alarms
```
